---
title: "leaflet"
author: "Henry Thorpe"
date: "4/7/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load libraries

```{r}
rm(list = ls())

for (pkg in c("tidyverse", "readr", "dplyr", "countrycode")) {library(pkg, character.only = TRUE)}
```
# Load data

```{r}
vaccine <- read_csv("data/vaccinations.csv")
#name_list <- c("OWID_ENG", "OWID_NIR", "OWID_SCT", "OWID_WLS")
#vaccine$iso_code <- with(vaccine, replace(iso_code, iso_code %in% name_list, "GBR"))
```
```{r}
coronavirus_summary <- read_csv("data/coronavirus_summary.csv")
coronavirus_summary$iso_code <- countrycode(coronavirus_summary$country, 'country.name','iso3c')
```
## Countries in the vaccine data not in summary data

```{r}
#subset(vaccine_by_country,duplicated(vaccine_by_country))
#subset(vaccine_by_country,duplicated(vaccine_by_country$country))
```


```{r}
unique(vaccine$country[!(vaccine$country %in% coronavirus_summary$country)])
```

## iso_code in the vaccine data not in summary data

```{r}
unique(vaccine$iso_code[!(vaccine$iso_code %in% coronavirus_summary$iso_code)])
```

## different iso_code correspondent country name in vaccine data

```{r}
unique(vaccine$country[vaccine$iso_code %in% c("OWID_ENG", "GGY", "JEY", "OWID_CYN", "OWID_NIR", "SHN", "OWID_SCT","OWID_WLS")])
```

# Check missing value

```{r}
apply(vaccine, 2, function(x) sum(is.na(x)))

```
```{r}
subset(vaccine,country %in% c("England", "United Kingdom"))
```


```{r}
str(vaccine)
```


# Aggregate by country

```{r}
vaccine_by_country<-vaccine%>%select(country,iso_code,,date,people_vaccinated,total_vaccinations,total_vaccinations_per_hundred,people_vaccinated_per_hundred,vaccines)%>%group_by(country)%>%filter(!is.na(total_vaccinations))%>%filter(total_vaccinations==max(total_vaccinations)) %>% slice(which.max(as.Date(date)))
head(vaccine_by_country, n=50)
```
## Total vaccinated by country

```{r,fig.height=20,fig.width=12}
options(repr.plot.width=30,repr.plot.height=20)
total_vaccinated<-vaccine_by_country%>%filter(!is.na(people_vaccinated)) %>% select(country,people_vaccinated)%>% arrange(desc(people_vaccinated))
total_vaccinated %>% ggplot()+geom_col(mapping = aes(x = reorder(country,people_vaccinated),y =people_vaccinated/10000000, fill=country), show.legend = FALSE)+
                    coord_flip()+
                theme_classic()+
                    ggtitle(label = "Total vaccinated  by countries.")+
                        xlab(label = "Countries")+
                        ylab(label = "millions")
```
##Top 10 countries with most vaccinated population

```{r}
options(repr.plot.width=30,repr.plot.height=20)
total_vaccinated<-vaccine_by_country%>%filter(!is.na(people_vaccinated)) %>% select(country,people_vaccinated)%>% arrange(desc(people_vaccinated))
total_vaccinated %>% head(10) %>% ggplot()+geom_col(mapping = aes(x = reorder(country,people_vaccinated),y =people_vaccinated/10000000, fill=country), show.legend = FALSE)+
                    coord_flip()+
                theme_classic()+
                    ggtitle(label = "Total vaccinated  by countries.")+
                        xlab(label = "Countries")+
                        ylab(label = "millions")
```
##Percent of total population vaccinated by country


```{r,fig.height=20,fig.width=12}
vaccine_prop_by_country<-vaccine_by_country %>% filter(!is.na(people_vaccinated_per_hundred)) %>% select(country,people_vaccinated_per_hundred) %>% arrange(desc(people_vaccinated_per_hundred))

vaccine_prop_by_country %>% arrange(desc(people_vaccinated_per_hundred)) %>%  ggplot()+geom_col(mapping = aes(x= reorder(country,people_vaccinated_per_hundred),y = people_vaccinated_per_hundred),fill="green")+
                    coord_flip()+
                        theme_light()+
                            ggtitle(label = "Proportion of the population vaccinated.")+
                                geom_text(aes(x = country,y =people_vaccinated_per_hundred,label=people_vaccinated_per_hundred ),hjust=-0.5)+
                                xlab(label = "Countries ")+
                                ylab(label = "Percentage")+
                                scale_y_continuous(limits = c(0,90),breaks = c(0,20,40))
```
## Top 15 countries with the biggest percentage of people vaccinated

```{r}
vaccine_prop_by_country<-vaccine_by_country %>% filter(!is.na(people_vaccinated_per_hundred)) %>% select(country,people_vaccinated_per_hundred) %>% arrange(desc(people_vaccinated_per_hundred))

vaccine_prop_by_country %>% head(15)%>%arrange(desc(people_vaccinated_per_hundred)) %>%  ggplot()+geom_col(mapping = aes(x= reorder(country,people_vaccinated_per_hundred),y = people_vaccinated_per_hundred),fill="green")+
                    coord_flip()+
                        theme_light()+
                            ggtitle(label = "Proportion of the population vaccinated.")+
                                geom_text(aes(x = country,y =people_vaccinated_per_hundred,label=people_vaccinated_per_hundred ),hjust=-0.5)+
                                xlab(label = "Countries ")+
                                ylab(label = "Percentage")+
                                scale_y_continuous(limits = c(0,90),breaks = c(0,20,40))
```
## Difference between people_vaccinated and toal_vaccinations

Some vaccines maybe end of no use because of any reason.

```{r}
head(vaccine_by_country[c("country","people_vaccinated","total_vaccinations")],20)
```

```{r}
#vaccine_by_country$use_rate = vaccine_by_country$people_vaccinated/vaccine_by_country$total_vaccinations
```

### Countries with less than 70% of vaccines were in use

```{r}
vaccine_use_rate_by_country<-vaccine_by_country %>% filter(!is.na(use_rate)) %>% select(country,use_rate) %>% arrange(desc(use_rate))

vaccine_use_rate_by_country %>% subset(use_rate<0.7)%>% ggplot(aes(x=country,y=use_rate))+geom_bar(stat="identity", position=position_dodge(),fill= "#00AFBB")+theme_bw()+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  ggtitle("Vaccine in use rate")+
  xlab("Country")+ 
  ylab("Vaccine in use rate")
```
### Only 5 countries with less than 60% of vaccines were in use

```{r}
subset(vaccine_by_country,vaccine_by_country$use_rate<0.6)
```

#Aggregated by vaccine brand


```{r}
vaccine_by_brand<-vaccine%>%select(vaccines,people_vaccinated)%>%group_by(vaccines)%>%filter(!is.na(people_vaccinated))%>%filter(people_vaccinated==max(people_vaccinated)) 
vaccine_by_brand
```
```{r}
#countries_cnt_by_brand<-vaccine%>%select(vaccines,country)%>%group_by(vaccines)%>%tally()
```

## Vaccines most widely used (by country numbers)

```{r,fig.height=15,fig.width=15}
options(repr.plot.width=12,repr.plot.height=14)
vaccine_by_country%>%group_by(vaccines)%>%
    ggplot(aes(x = vaccines)) +
    geom_bar(aes(fill = country), show.legend = FALSE) +
    labs(x = "Vaccines", y = "Number Of Countries") +
    theme(axis.text.x = element_text(size=16,angle = 90, vjust = 0.5, hjust=1),
          axis.text.y = element_text(size=16),
          axis.title.x = element_text(size=20),
          axis.title.y = element_text(size=20))
```

## Vaccines most widely used (by number of people)

```{r,fig.height=15,fig.width=15}
ggplot(vaccine_by_brand,aes(x=vaccines,y=people_vaccinated)) + geom_bar(stat='identity',fill="blue", width=.5)+labs(x = "Vaccines", y = "Total number of people vaccinated") + theme_classic()+
    theme(axis.text.x = element_text(size=16,angle = 90, vjust = 0.5, hjust=1),
          axis.text.y = element_text(size=16),
          axis.title.x = element_text(size=20),
          axis.title.y = element_text(size=20))
```



