---
title: "Data Analysis 2"
author: "Alexis Cenname"
date: "9/30/2020"
output:
  html_document:
    df_print: paged
    toc: true
    number_sections: true
    toc_depth: '3'
    code_folding: show
---

```{r,echo=FALSE,message=FALSE,warning=FALSE}
require(knitr)
# Set so that long lines in R will be wrapped:
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE)
```

# Load Library

```{r}
library(ggplot2)
library(dygraphs)
library(countrycode)
library(dplyr)
```

# Data Import

```{r}
#crystal code

rm(list = ls())

for (pkg in c("tidyverse", "readr", "dplyr", "countrycode")) {library(pkg, character.only = TRUE)}

coronavirus_summary <- read.csv("C:/Users/Alexis Cenname/Desktop/Advanced R Computing/biost2094_project/data/coronavirus_summary.csv")
coronavirus_summary$iso_code <- countrycode(coronavirus_summary$country, 'country.name','iso3c')
coronavirus_summary <- select(coronavirus_summary, -country)
vaccine <- read.csv("C:/Users/Alexis Cenname/Desktop/Advanced R Computing/biost2094_project/data/vaccinations.csv")
name_list <- c("OWID_ENG", "OWID_NIR", "OWID_SCT", "OWID_WLS")
vaccine$iso_code <- with(vaccine, replace(iso_code, iso_code %in% name_list, "GBR"))
combine <- merge(vaccine, coronavirus_summary, by="iso_code", all.x=T)
coronavirus_daily <- read.csv("C:/Users/Alexis Cenname/Desktop/Advanced R Computing/biost2094_project/data/coronavirus_daily.csv")
```


#Top 20 total deaths from COVID by country

```{r}

#aggregate by country, continent, max total deaths

aggregated_deaths <- aggregate(combine$total_deaths, by = list(combine$country, combine$continent), max) %>% arrange(desc(x)) %>% drop_na(x)

#plot
ggplot(aggregated_deaths[1:20,], aes(x, y=Group.1, fill=Group.1)) + geom_col() + theme_classic() + labs(title="Top 20 Total Deaths From COVID By Country", y="Country", x="Total Deaths", fill="Country Legend") + theme(legend.position = "none")
```



#Top 20 recoveries by country

```{r}
aggregated_recoveries <- aggregate(combine$total_recovered, by = list(combine$country, combine$continent), max) %>% arrange(desc(x)) %>% drop_na(x)

ggplot(aggregated_recoveries[1:20,], aes(x, y=Group.1, fill=Group.1)) + geom_col() + theme_classic() + labs(title="Top 20 Total Recoveries By Country", y="Country", x="Total Recoveries", fill="Country Legend") + theme(legend.position = "none")
```

#Top 20 serious cases by country

```{r}
aggregated_serious <- aggregate(combine$serious_or_critical, by = list(combine$country, combine$continent), max) %>% arrange(desc(x)) %>% drop_na(x)

ggplot(aggregated_serious[1:20,], aes(x, y=Group.1, fill=Group.1)) + geom_col() + theme_classic() + labs(title="Top 20 Total Serious Cases By Country", y="Country", x="Total Serious/Critical Cases", fill="Country Legend") + theme(legend.position = "none")
```

#Top 20 Total Tests by country

```{r}
aggregated_tests <- aggregate(combine$total_tests, by = list(combine$country, combine$continent), max) %>% arrange(desc(x)) %>% drop_na(x)

ggplot(aggregated_tests[1:20,], aes(x/1e8, y=Group.1, fill=Group.1)) + geom_col() + theme_classic() + labs(title="Top 20 Total Tests By Country", y="Country", x="Total Tests (*1e8)", fill="Country Legend") + theme(legend.position = "none")
```


#Top 20 Total Deaths Per One Million by country

```{r}
aggregated_death_per_mil <- aggregate(combine$total_deaths_per_1m_population, by = list(combine$country, combine$continent), max) %>% arrange(desc(x)) %>% drop_na(x)

ggplot(aggregated_death_per_mil[1:20,], aes(x, y=Group.1, fill=Group.1)) + geom_col() + theme_classic() + labs(title="Top 20 Deaths Per Million By Country", y="Country", x="Total Deaths Per Million", fill="Country Legend") +theme(legend.position = "none")
```


#Top 20 Total Tests Per Million

```{r}
aggregated_test_per_mil <- aggregate(combine$total_tests_per_1m_population, by = list(combine$country, combine$continent), max) %>% arrange(desc(x)) %>% drop_na(x)

ggplot(aggregated_test_per_mil[1:20,], aes(x, y=Group.1, fill=Group.1)) + geom_col() + theme_classic() + labs(title="Top 20 Tests Per Million By Country", y="Country", x="Total Tests", fill="Country Legend") + theme(legend.position = "none")
```


#Total Deaths By Continent

```{r}
aggregate_death_continent <- aggregate(aggregated_deaths$x, by = list(aggregated_deaths$Group.2), sum) %>% drop_na(x)

ggplot(aggregate_death_continent, aes(x/1e5, y=Group.1, fill=Group.1)) + geom_col() + theme_classic() + labs(title="Total Deaths By Continent", y="Continent", x="Total Deaths(*1e5)", fill="Continent Legend") + coord_flip() + theme(legend.position = "none")
```


#Total Recovered by Continent

```{r}
aggregate_recovered_continent <- aggregate(aggregated_recoveries$x, 
                                           by = list(aggregated_recoveries$Group.2), sum) %>% 
                                           drop_na(x)

ggplot(aggregate_recovered_continent, aes(log(x), y=Group.1, fill=Group.1)) + 
       geom_col() + 
       theme_classic() + 
       labs(title="Total Recoveries By Continent", x="log(Total Recovered)", fill="Continent Legend") + 
       coord_flip() + 
       theme(axis.text.x=element_blank(), axis.title.x = element_blank())
```



# Merge vaccine and daily_coronavirus

```{r}
vac_daily <- merge(coronavirus_daily, vaccine, by=c("country", "date"))
```


# Active cases vs. total vaccinated (Afghanistan)

```{r}
afghanistan <- vac_daily %>% 
               drop_na(active_cases, date) %>% 
               subset(country=="Afghanistan")

ggplot(afghanistan, aes(date, active_cases)) + 
       geom_point(color="pink") + 
       geom_line(group=1, color="pink") + 
       labs(title="Active Cases Daily", x="Dates", y="Active Cases") + 
       theme_classic() + 
       theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), 
             legend.position = "none")

```

# Deaths per day (Afghanistan)

```{r}
afghanistan_2 <- vac_daily %>% 
                 drop_na(daily_new_deaths, date) %>% 
                 subset(country=="Afghanistan")

ggplot(afghanistan_2, aes(date, daily_new_deaths)) + 
       geom_point(color="black") + geom_line(group=1) + 
       labs(title="New Deaths Daily", x="Dates", y="Deaths") + 
       theme_classic() + 
       theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), 
             legend.position = "none")
```

# Libraries for RShiny

```{r}
library(readr)
library(magrittr)
library(rvest)
library(readxl)
library(dplyr)
library(DT)
library(maps)
library(ggplot2)
library(reshape2)
library(ggiraph)
library(RColorBrewer)
library(leaflet)
library(plotly)
library(geojsonio)
library(shiny)
library(shinyWidgets)
library(shinydashboard)
library(shinythemes)
library(mapproj)
library(janitor)
```

```{r}
rm(list = ls())

for (pkg in c("tidyverse", "readr", "dplyr", "countrycode")) {library(pkg, character.only = TRUE)}

coronavirus_summary <- read.csv("C:/Users/Alexis Cenname/Desktop/Advanced R Computing/biost2094_project/data/coronavirus_summary.csv")
coronavirus_summary$iso_code <- countrycode(coronavirus_summary$country, 'country.name','iso3c')
coronavirus_summary <- select(coronavirus_summary, -country)
vaccine <- read.csv("C:/Users/Alexis Cenname/Desktop/Advanced R Computing/biost2094_project/data/vaccinations.csv")
name_list <- c("OWID_ENG", "OWID_NIR", "OWID_SCT", "OWID_WLS")
vaccine$iso_code <- with(vaccine, replace(iso_code, iso_code %in% name_list, "GBR"))
combine <- merge(vaccine, coronavirus_summary, by="iso_code", all.x=T)
coronavirus_daily <- read.csv("C:/Users/Alexis Cenname/Desktop/Advanced R Computing/biost2094_project/data/coronavirus_daily.csv")
vac_daily <- merge(coronavirus_daily, vaccine, by=c("country", "date"))

ui <- navbarPage(title = "COVID-19 Vaccine",
                 # First Page
                 tabPanel(title = "About the site",
                          tags$br(),tags$br(),tags$h4("Background"),
                          "SARS-CoV-2 has impacted the world in an unprecedented way,
                      with the world having changed significantly since H2N2/H3N2
                      pandemics in the mid 1900’s and even more since the major comparison
                      point, the 1918 flu pandemic. The world is much more global now.
                      Adjusted for inflation, the value of goods shipped internationally in
                      2014 was six times higher than the value of shipped goods in 1969, the
                      last year of the H3N2 pandemic. In comparison to 1919, that number goes
                      all the way to 53 times as much as shipped in 2014 (Beltekian). As such,
                      returning to normalcy is a worldwide goal.",
                          tags$br(),tags$br(),
                          "Right now, the main indicator
                      we have is the vaccination rate for each country, each of which has its
                      own unique situation and challenges to account for, so viewing that data
                      quickly and clearly is an important piece to know how close we are to the
                      end of the pandemic.",

                          tags$br(),tags$br(),tags$h4("Code"),
                          "Code and input data used to generate this Shiny mapping tool are available on ",tags$a(href="https://github.com/czang97/biost2094_project", "Github."),

                          tags$br(),tags$br(),tags$h4("Contributor"),
                          br(),
                          "Crystal Zang",
                          br(),
                          "Liling Liu",
                          br(),
                          "Alexis Cenname",
                          br(),
                          "Henry Thrope",
                          br(),
                          br(),
                          br(),
                          br(),
                          br(),
                          img(src = "logo.jpeg",height = 130, width=250)),
                 tabPanel(title = "Worldwide Vaccine Progress"),
                 tabPanel(title = "Vaccine Progress Map"),
                 tabPanel(title = "US Vaccine Progress"),
                 tabPanel(title = "Reaching Herd Immunity",
                          sidebarLayout(
                          sidebarPanel(

                              # Options to select population
                              selectInput("countrylist",
                                            label = h4("Country"),
                                            choices = vac_daily %>% drop_na(date, active_cases) %>% select(country),
                                            selected = "Afghanistan")),
                          mainPanel(plotOutput("countryplot"), plotOutput("countryplot2"), plotOutput("countryplot3"))
                          )
                          ))

server <- function(input, output) {

  output$countryplot <- renderPlot(ggplot(vac_daily %>% filter(country==input$countrylist) %>% 
                                          drop_na(active_cases), 
                                          aes(date, active_cases))
                                  + geom_point(color="red", size=5) + geom_line(group=1, color="black") + theme_bw() +
                                     theme(axis.text.x = element_text(angle = 90, vjust =0.5, hjust=1), legend.position = "none",
                                           axis.title.x = element_blank()) +
                                     labs(title = as.character(input$countrylist), y = "Active Cases"))
  output$countryplot2 <- renderPlot(ggplot(vac_daily %>% filter(country==input$countrylist) %>% 
                                          drop_na(daily_vaccinations), 
                                          aes(date, daily_vaccinations))
                                  + geom_point(color="green", size=5) + geom_line(group=1, color="black") + theme_bw() +
                                     theme(axis.text.x = element_text(angle = 90, vjust =0.5, hjust=1), legend.position = "none",
                                           axis.title.x = element_blank()) +
                                     labs(y = "Daily Vaccinations"))
  output$countryplot3 <- renderPlot(ggplot(vac_daily %>% filter(country==input$countrylist) %>% 
                                          drop_na(total_vaccinations), 
                                          aes(date, total_vaccinations))
                                  + geom_point(color="purple", size=5) + geom_line(group=1, color="black") + theme_bw() +
                                     theme(axis.text.x = element_text(angle = 90, vjust =0.5, hjust=1), legend.position = "none",
                                           axis.title.x = element_blank()) +
                                     labs(y = "Total Vaccinations")
  )
}

# Run the app ----
shinyApp(ui = ui, server = server)
```





