---
title: "Data Analysis 2"
author: "Alexis Cenname"
date: "9/30/2020"
output:
  html_document:
    df_print: paged
    toc: true
    number_sections: true
    toc_depth: '3'
    code_folding: show
---

```{r,echo=FALSE,message=FALSE,warning=FALSE}
require(knitr)
# Set so that long lines in R will be wrapped:
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE)
```

# Load Library

```{r}
library(ggplot2)
library(dygraphs)
library(countrycode)
library(dplyr)
```

# Data Import

```{r}
#crystal code

rm(list = ls())

for (pkg in c("tidyverse", "readr", "dplyr", "countrycode")) {library(pkg, character.only = TRUE)}

coronavirus_summary <- read.csv("C:/Users/Alexis Cenname/Desktop/Advanced R Computing/biost2094_project/data/coronavirus_summary.csv")
coronavirus_summary$iso_code <- countrycode(coronavirus_summary$country, 'country.name','iso3c')
coronavirus_summary <- select(coronavirus_summary, -country)
vaccine <- read.csv("C:/Users/Alexis Cenname/Desktop/Advanced R Computing/biost2094_project/data/vaccinations.csv")
name_list <- c("OWID_ENG", "OWID_NIR", "OWID_SCT", "OWID_WLS")
vaccine$iso_code <- with(vaccine, replace(iso_code, iso_code %in% name_list, "GBR"))
combine <- merge(vaccine, coronavirus_summary, by="iso_code", all.x=T)
coronavirus_daily <- read.csv("C:/Users/Alexis Cenname/Desktop/Advanced R Computing/biost2094_project/data/coronavirus_daily.csv")
```


#Top 20 total deaths from COVID by country

```{r}

#aggregate by country, continent, max total deaths

aggregated_deaths <- aggregate(combine$total_deaths, by = list(combine$country, combine$continent), max) %>% arrange(desc(x)) %>% drop_na(x)

#plot
ggplot(aggregated_deaths[1:20,], aes(x, y=Group.1, fill=Group.1)) + geom_col() + theme_classic() + labs(title="Top 20 Total Deaths From COVID By Country", y="Country", x="Total Deaths", fill="Country Legend") + theme(legend.position = "none")
```



#Top 20 recoveries by country

```{r}
aggregated_recoveries <- aggregate(combine$total_recovered, by = list(combine$country, combine$continent), max) %>% arrange(desc(x)) %>% drop_na(x)

ggplot(aggregated_recoveries[1:20,], aes(x, y=Group.1, fill=Group.1)) + geom_col() + theme_classic() + labs(title="Top 20 Total Recoveries By Country", y="Country", x="Total Recoveries", fill="Country Legend") + theme(legend.position = "none")
```

#Top 20 serious cases by country

```{r}
aggregated_serious <- aggregate(combine$serious_or_critical, by = list(combine$country, combine$continent), max) %>% arrange(desc(x)) %>% drop_na(x)

ggplot(aggregated_serious[1:20,], aes(x, y=Group.1, fill=Group.1)) + geom_col() + theme_classic() + labs(title="Top 20 Total Serious Cases By Country", y="Country", x="Total Serious/Critical Cases", fill="Country Legend") + theme(legend.position = "none")
```

#Top 20 Total Tests by country

```{r}
aggregated_tests <- aggregate(combine$total_tests, by = list(combine$country, combine$continent), max) %>% arrange(desc(x)) %>% drop_na(x)

ggplot(aggregated_tests[1:20,], aes(x/1e8, y=Group.1, fill=Group.1)) + geom_col() + theme_classic() + labs(title="Top 20 Total Tests By Country", y="Country", x="Total Tests (*1e8)", fill="Country Legend") + theme(legend.position = "none")
```


#Top 20 Total Deaths Per One Million by country

```{r}
aggregated_death_per_mil <- aggregate(combine$total_deaths_per_1m_population, by = list(combine$country, combine$continent), max) %>% arrange(desc(x)) %>% drop_na(x)

ggplot(aggregated_death_per_mil[1:20,], aes(x, y=Group.1, fill=Group.1)) + geom_col() + theme_classic() + labs(title="Top 20 Deaths Per Million By Country", y="Country", x="Total Deaths Per Million", fill="Country Legend") +theme(legend.position = "none")
```


#Top 20 Total Tests Per Million

```{r}
aggregated_test_per_mil <- aggregate(combine$total_tests_per_1m_population, by = list(combine$country, combine$continent), max) %>% arrange(desc(x)) %>% drop_na(x)

ggplot(aggregated_test_per_mil[1:20,], aes(x, y=Group.1, fill=Group.1)) + geom_col() + theme_classic() + labs(title="Top 20 Tests Per Million By Country", y="Country", x="Total Tests", fill="Country Legend") + theme(legend.position = "none")
```


#Total Deaths By Continent

```{r}
aggregate_death_continent <- aggregate(aggregated_deaths$x, by = list(aggregated_deaths$Group.2), sum) %>% drop_na(x)

ggplot(aggregate_death_continent, aes(x/1e5, y=Group.1, fill=Group.1)) + geom_col() + theme_classic() + labs(title="Total Deaths By Continent", y="Continent", x="Total Deaths(*1e5)", fill="Continent Legend") + coord_flip() + theme(legend.position = "none")
```


#Total Recovered by Continent

```{r}
aggregate_recovered_continent <- aggregate(aggregated_recoveries$x, by = list(aggregated_recoveries$Group.2), sum) %>% drop_na(x)

ggplot(aggregate_recovered_continent, aes(x/1e5, y=Group.1, fill=Group.1)) + geom_col() + theme_classic() + labs(title="Total Recoveries By Continent", y="Continent", x="Total Recovered(*1e6)", fill="Continent Legend") + coord_flip()
```



# Merge vaccine and daily_coronavirus

```{r}
vac_daily <- merge(coronavirus_daily, vaccine, by=c("country", "date"))
```


# Active cases vs. total vaccinated (Afghanistan)

```{r}
afghanistan <- vac_daily %>% drop_na(active_cases, date) %>% subset(country==c("Afghanistan"))
ggplot(afghanistan, aes(date, active_cases, color="pink")) + geom_point() + geom_line(group=1) + labs(title="Active Cases Daily", y="Dates", x="Active Cases") + theme_classic() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + theme(legend.position = "none")

```

# Deaths per day (Afghanistan)

```{r}
afghanistan_2 <- vac_daily %>% drop_na(daily_new_deaths, date) %>% subset(country==c("Afghanistan"))
ggplot(afghanistan_2, aes(date, daily_new_deaths, color="black")) + geom_point() + geom_line(group=1) + labs(title="New Deaths Daily", y="Dates", x="Deaths") + theme_classic() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + theme(legend.position = "none")
```

# Create tab in RShiny

```{r}

```

